---
title: 'NOAA Chinook and Chum salmon mixed stock analysis'
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
author: "Patrick Dylan Barry"
date: "15 April 2019"
output:
  html_document:
    df_print: paged
bibliography: ./Bibliography/NOAA_MSA.bib
---

This file is a slightly modified version of Patrick's original one. 
Primary changes include:
1) Removes creation of the chum baseline file that includes genotypes instead of just allele frequencies.
2) Puts the functions into a loop so that it will run rubias for all of the .mix files that are in the ./Data/Chum/mixture_files folder.
3) Puts the functions into a loop so that it will compare rubias outputs with all of the Bayes outputs in the ./Data/Chum/Bayes/2018 folder.
4) Figures of the Bayes/rubias comparison are output into the ./Data/Chum/Bayes_Rubias_Compare_Figs folder.

```{r setup, include = FALSE}
# !diagnostics off
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = TRUE
)
```


```{r loadpackages, include = FALSE}
library("stringr")
library("tidyverse")
library("rubias")
```

Define globally relevant terms and functions
```{r createGlobalVals,eval=TRUE}
myfiles <- gsub(pattern = "\\.mix$", "",list.files("Data/Chum/mixture_files"))
mycsv <- list.files("./Data/Chum/rubias/2018/")
#now lets deal with the mixture
#This should be easy with the script from the King salmon
source("./functions/Bayes2Genepop.R")
source("./functions/Genepop2rubias.R")

RGnames<-matrix(data=c(seq(1,6,1),c("EAsia","NAsia","WAlaska","UpMidYukon","SW_Alaska","E_GOA_PNW")),nrow=6,ncol=2)
nloci <- readLines("./Data/Chum/Loci.txt") %>% length()
Convert2uSat<-read.csv("./Data/Chum/ConvertuSats.csv")

cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
myburn <- 5000
modelreps <- 10000
```

Convert Bayes mixture files, first to genepop and then from genepop to rubias formats.
Output all of the .gen files into a folder called ./Data/Chum/gen_files.
Output all of the .csv files into a folder called ./Data/Chum/rubias/2018

```{r createRubiasMix,eval=FALSE}
#Bayes2Genepop_mixture(infile="./Data/Chum/Age3Area1.MIX",outfile="./Data/Chum/Age3Area1.gen",LociFile="./Data/Chum/Loci.txt",digits=3,SampPrefix="Mix",Project="Age3Area1_Bayes2Genepop")
#Genepop2rubias_mixture(infile="./Data/Chum/Age3Area1.gen",outfile="./Data/Chum/rubias/Age3Area1.csv",digits=3)

lapply(myfiles,function(x) Bayes2Genepop_mixture(infile=paste0("./Data/Chum/mixture_files/",x,".mix"),
                                                      outfile=paste0("./Data/Chum/gen_files/",x,".gen"),
                                                      LociFile="./Data/Chum/Loci.txt",
                                                      digits=3,
                                                      SampPrefix="Mix",
                                                      Project=paste0(x,"_Bayes2Genepop")))

lapply(myfiles,function(x) Genepop2rubias_mixture(infile=paste0("./Data/Chum/gen_files/",x,".gen"),
                                                  outfile=paste0("./Data/Chum/rubias/2018/",x,".csv"),
                                                  digits=3))

#Our alleles in this file are the index number from the baseline, we need to convert to uSat allele size
mycsv <- list.files("./Data/Chum/rubias/2018/")

for(i in 1:length(mycsv)){
  rubiasMix<-read.csv(paste0("./Data/Chum/rubias/2018/",mycsv[i]))
  rubiasMix<-as.matrix(rubiasMix)
  for (l in 1:nloci){
    rubiasMix[,(l*2+3):(l*2+4)]<-plyr::mapvalues(x=str_trim(rubiasMix[,(l*2+3):(l*2+4)]),from=Convert2uSat[which(Convert2uSat[,1]==l),4],to=Convert2uSat[which(Convert2uSat[,1]==l),3],warn_missing = F)
  }
  
  write.csv(x=rubiasMix,paste0("./Data/Chum/rubias/2018/",mycsv[i]),row.names = F)
}

```

Run the rubias model, using the model iterations and burn-in specifed in the global values chunk at the beginning of this document.

```{r ChumRubiasBayes, eval=FALSE, cache.lazy=FALSE}
#Run rubias from within R
for(i in 1:length(mycsv)){
myout<-read.csv(paste0("./Data/Chum/rubias/2018/",mycsv[i]),stringsAsFactors = F)
ChumBase<-read.csv("./Data/Chum/rubias/ChumBaseline_rubias.csv",stringsAsFactors = F)

infer_mixture(reference = ChumBase, 
                         mixture = myout, 
                         gen_start_col = 5,reps = modelreps, burn_in = myburn) %>% 
  saveRDS(.,file=paste0("./Data/Chum/",myfiles[i],".RDS"))
}
```

Create summaries from the rubias files and plot (and save) comparisons for each of the GSI strata (e.g., cluster, season, etc.)

```{r}

for(i in 1:length(myfiles)){
  #  Load rubias output
  mix_est <- readRDS(paste0("./Data/Chum/",myfiles[i],".RDS"))
  
  #clean up the results
  RG_mix_ests <- mix_est$mixing_proportions %>%
    dplyr::group_by(mixture_collection, repunit) %>%
    dplyr::summarise(repprop = sum(pi)) 
  
  trace_subset <- mix_est$mix_prop_traces %>%
    dplyr::filter(mixture_collection == "Mix", sweep > myburn) %>%
    dplyr::group_by(sweep, repunit) %>%
    dplyr::summarise(repprop = sum(pi)) 
  
  CI_rub <- trace_subset %>%
    dplyr::group_by(repunit) %>%
    dplyr::summarise(loCI = quantile(repprop, probs = 0.025),
                     hiCI = quantile(repprop, probs = 0.975))
  
  #RubiasRes<-cbind(RG_mix_ests[,2:3],CI_rub[,2:3])
  #RubiasRes$InfMeth<-"Rubias"
  
  RubiasRes <- RG_mix_ests %>% 
    inner_join(CI_rub) %>% 
    ungroup %>% 
    dplyr::select(-mixture_collection) %>% 
    mutate(InfMeth="Rubias")
  
  #Pull in the Bayes results from .SUM files
  Res<-readLines(paste0("./Data/Chum/Bayes/2018/sum_files/",myfiles[i],"_Chain1_estimate.SUM"))
  StockNum<-6
  StockProp_Bayes<-Res[(grep("Chains combined: ",Res)+2):(grep("Chains combined: ",Res)+3+StockNum)]
  
  #  Create a matrix of results, tidy, bind together rubias and bayes outputs
  Res <- matrix(data=unlist(stringr::str_split(StockProp_Bayes[-(1:2)],"\\s+")),nrow=StockNum,ncol=9,byrow = T)[,-c(1:2)] %>% 
    data.frame %>% 
    dplyr::select(5,4,6) %>% 
    mutate_all(list(~as.numeric(as.character(.)))) %>% 
    mutate(repunit=RGnames[,2],
           InfMeth="Bayes") %>% 
    dplyr::select(repunit,everything()) %>% 
    setNames(colnames(RubiasRes)) %>% 
    bind_rows(RubiasRes)
  
  # The following commented out code is not merged into the previous "res" chunk.
  #create a matrix of the results
  #StockPropMat<-matrix(data=unlist(stringr::str_split(StockProp_Bayes[-(1:2)],"\\s+")),nrow=StockNum,ncol=9,byrow = T)[,-c(1:2)]
  
  #tidy up the results
  #StockPropMat<-StockPropMat[,c(1,5,4,6)]
  #map the reporting group names instead of the numbers
  #StockPropMat[,1] <- RGnames[,2]
  #StockPropMat[,1]<-plyr::mapvalues(as.numeric(StockPropMat[,1]),from=RGnames[,1],to=as.character(RGnames[,2]))
  
  #create a matrix that we can bind with the rubias results for graphing
  #BayesRes<-cbind(StockPropMat,rep("Bayes",nrow(StockPropMat)))
  #colnames(BayesRes)<-colnames(RubiasRes)
  
  #bind up all the results
  #Res<-rbind(RubiasRes,BayesRes)
  #Res[,2]<-as.numeric(Res[,2])
  #Res[,3]<-as.numeric(Res[,3])
  #Res[,4]<-as.numeric(Res[,4])
  
  #rearrange the factor level so that the regions are in the order from the report
  #Res[,1]<-factor(Res[,1],levels(as.factor(Res[,1]))[c(2,3,6,5,4,1)])
  
  #  Specify the plotting order
  myorder <- c("SE Asia","NE Asia","Western AK","Upper/Middle Yukon","SW Alaska","Eastern GOA/PNW")

  png(paste0("./Data/Chum/Bayes_Rubias_Compare_Figs/",myfiles[i],"_median.png"),width=6,height=6,units="in",res=300)
  print(Res %>% 
          mutate(repunit=fct_recode(repunit,
                                    "SE Asia"="EAsia",
                                    "NE Asia"="NAsia",
                                    "Western AK"="WAlaska",
                                    "Upper/Middle Yukon"="UpMidYukon",
                                    "SW Alaska"="SW_Alaska",
                                    "Eastern GOA/PNW"="E_GOA_PNW"),
                 repunit2=fct_relevel(repunit,myorder)) %>% 
    ggplot(aes(x=repunit2, y=repprop, fill=InfMeth)) + 
    geom_bar(stat="identity", color="black", 
             position=position_dodge()) +
    geom_errorbar(aes(ymin=loCI, ymax=hiCI), width=.2,
                  position=position_dodge(.9)) + 
    theme_bw() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))+
    scale_fill_manual(values=cbPalette[2:3])+
    guides(fill=guide_legend(title="Inference \nMethod"))+
    xlab("Reporting Group")+
    ylab("Stock Composition") + 
    ggtitle(paste0(myfiles[i])))
  dev.off()
}
```
